Timbral similarity between extended playing techniques: the next frontier in querying instrumental databases

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
"i.e." need not be italicized (IEEE). Search and replace

\newcommand*{\eg}{e.g.\@\xspace}
\newcommand{\ipt}{IPT\xspace}
\newcommand{\ipts}{IPTs\xspace}

@@

Interestingly, modern physiological models of the primary mammalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.  Dau's model accounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}. Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}. Though, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefulness of the bandwidth dimension \cite{mesgarani2006discrimination}.

@@

Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider the relation between the rate of modulation and timbre percepts. A notion that have not been extensively studied in the literature \cite{burred2010dynamic}. Compared to the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at a complementary finer grain of detail.

@@

Studies about perceptual similarity are plagued with a dimensional problem.
When considering $n$ items, a complete exploration of the similarity space
requires the filling of a $n^2$ matrix. Assuming a symmetric similarity,
\textit{i.e.} the similarity of A to B is equal to the similarity of B to A,
reduces the number but not by much.

@@

Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \textit{i.e.} a very diverse set of instruments labels for the \ipt they belong to. Also, some clusters could be merged without stong changes in the overall organisation. This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.

>>>>

\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\newcommand{\ipt}{IPT\xspace}
\newcommand{\ipts}{IPTs\xspace}

@@

Interestingly, modern physiological models of the primary mammalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \ie{} how the energy modulates through time for given frequencies.  Dau's model accounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}. Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}. Though, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefulness of the bandwidth dimension \cite{mesgarani2006discrimination}.

@@

Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider the relation between the rate of modulation and timbre percepts. A notion that have not been extensively studied in the literature \cite{burred2010dynamic}. Compared to the reference that is usually taken in timbre studies \ie{} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at a complementary finer grain of detail.

@@

Studies about perceptual similarity are plagued with a dimensional problem.
When considering $n$ items, a complete exploration of the similarity space
requires the filling of a $n^2$ matrix. Assuming a symmetric similarity,
\ie{} the similarity of A to B is equal to the similarity of B to A,
reduces the number but not by much.

@@

Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to. Also, some clusters could be merged without stong changes in the overall organisation. This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
